### 2022-02-09 symposium.tex
Type and Null Safety for Regexes

\emph{In principle}, Scala~3's type system is capable of doing a lightweight form of dependent typing (\Cref{subsec:a-lightweight-form-of-dependent-typing}), where it...

\begin{lstlisting}
def isMarked
  [R <: String, At <: Int, Hi <: Int]
  (r: R, at: At, hi: Hi)
  : IsMarked[R, At, Hi]
\end{lstlisting}

>>>>

This solution has the benefit of being conceptually simple.
We needed to write an implementation that conforms to our type-level program, and we did!
By duplicating that program, and using an unsafe case to convince the compiler of our good intentions.

The first alternative that comes to mind is perhaps to turn ourselves to a dependently typed language that supports type- and term-level polymorphism.
In a language with shared term and type syntax, we could write our analysis generically and derive two programs from it, one for each level.

In principle, Scala~3's type inference should be able to solve half of the problem with its ability to correlate match types and match expression.
In \Cref{subsec:a-lightweight-form-of-dependent-typing}, we show an example of that ability, where the compiler type checks a small program using a match type.
In our case, this mechanism should allow us to get rid of the unsafe cast but does not address the problem of code duplication.
In practice, at the time of writing, this solution does not play well with predefined types from the standard library, because the compiler lacks the knowledge necessary to correlate those types with their term-level counterpart.
For instance, the compiler does not recognize |CharAt["foo",0]| as a valid type for |"foo".chatAt(0)|.

In the long run, it would be interesting to investigate extending Scala with a "precise" mode of type inference, that would automatically infer match types and singleton types, whenever possible.
In \Cref{chap:generalizing-singleton-types}, we demonstrated the feasibility of this approach by generalizing Scala's singleton types.
A promising avenue would be to revisit that line of work through the lens of type inference.
